### 2022-11-18 chapter1/chapter1.tex
Added discussion of setting local k values

For example, despite many additional pre-processing steps and other tweaks introduced by later authors, the basic framework of orthology inference by clustering the hit graph has remained largely unchanged in the past twenty years~\cite{Tatusov1997, Remm2001, Li2003, Jensen2007, Linard2011, Emms2015, Train2017, Cosentino2018}. This longevity is a testament to the robustness of the underlying idea that orthologous proteins should consistently identify each other as the most similar pairs between their genomes. Even as the number of genomes and their taxonomic density has increased dramatically, many orthology inference pipelines continue to use algorithms which were originally applied to sets of far fewer and more distantly related genomes. This mismatch in scale increases the chance of propagating annotation errors since only a small number of edges are needed to create or merge clusters. Thus, we instead applied a generalization of the triangle and connected components clustering methods called \textit{k}-clique percolation where \textit{k} is a tunable parameter that influences the tightness of a cluster. The optimal value of \textit{k} for a given set of genomes is unclear and likely depends on the desired trade-off between sensitivity and specificity. However, percolation theory or simulations may yield insights.

>>>>

For example, despite many additional pre-processing steps and other tweaks introduced by later authors, the basic framework of orthology inference by clustering the hit graph has remained largely unchanged in the past twenty years~\cite{Tatusov1997, Remm2001, Li2003, Jensen2007, Linard2011, Emms2015, Train2017, Cosentino2018}. This longevity is a testament to the robustness of the underlying idea that orthologous proteins should consistently identify each other as the most similar pairs between their genomes. Even as the number of genomes and their taxonomic density has increased dramatically, many orthology inference pipelines continue to use algorithms which were originally applied to sets of far fewer and more distantly related genomes. This mismatch in scale increases the chance of propagating annotation errors since only a small number of edges are needed to create or merge clusters. Thus, we instead applied a generalization of the triangle and connected components clustering methods called \textit{k}-clique percolation where \textit{k} is a tunable parameter that influences the tightness of a cluster. The optimal value of \textit{k} for a given set of genomes is unclear and likely depends on the desired trade-off between sensitivity and specificity. Furthermore, \textit{k} is not necessarily a global parameter and can instead depend on the properties of each connected component. For example, one possibility is to take an entire component as an orthologous group if its number of unique genes and unique species are equal since all the sequences are isoforms of a single set of genes. This would effectively set \text{k} equal to one for this component. Another approach is to make \textit{k} an decreasing function of the density of edges, so sparser graphs are clustered more permissively. However, percolation theory or simulations may yield additional insights.
